Forecast of the 2024 Presidential Election through the Lens of Turbulent Swing States

### 2024-11-04 paper/paper.qmd

---
title: "Forecast of the 2024 Presidential Election through the Lens of Turbulent Swing States"
subtitle: "Kamala Harris Leads the Race With 276 Electoral Votes With Support in 4 of 7 Swing States"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/president-polls)."
date: today
date-format: long
abstract: "The 2024 U.S. presidential election is shaping up to be a highly swing race, with polling data playing a crucial role in gauging public sentiment and forecasting the potential outcome. This paper builds a predictive model using a poll-of-polls approach, aggregating data from multiple pollsters to forecast the winner of the upcoming election. The analysis incorporates polling data from various sources, filtered to include only high-quality pollsters with a numeric grade above 3, to ensure the reliability of the predictions. By incorporating the poll-of-polls approach with our model, we predict that Kamla Harris edges our Trump in the presidential election. Our model forecasts that with a 1%-4% margin lead in Pennsylvania, Wisconsin, Nevada, and Michigan, Kamala Harris will accumulate 276 electoral votes to win the election. Additionally, a detailed examination of one selected pollster’s methodology is conducted, highlighting strengths, weaknesses, and biases in the survey process. The paper concludes with an idealized polling methodology, designed within a $100K budget, to enhance the accuracy of election forecasts. Results are presented in the context of understanding polling trends, potential biases, and the evolving landscape of voter preferences as the election approaches."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Appendix A {-}

#  Sampling Methodology Employed by New York Times and Siena College:

@@

## Target Population and Sampling Frame
The poll targets registered voters in the United States, with a particular focus on voters in swing states for presidential races. Consequently, the target population is registered voters of the United States of America and the sampling frame is the voter file, a national list of registered voters from which NYC employs random sampling to create their pollsters [@nytpoll]. This voter file contains information on more than 200 million Americans that includes information about their party affiliation, voting history, and  basic demographic information. From this sampling frame, respondents are randomly selected while balancing key demographic and political characteristics through both front-end sampling and post-survey weighting pursuant to appropriate representation of the entire target population [@nytpoll].

## Recruitment Methods
The poll relies on random digit dialing to reach voters, predominantly via cell phones. While landlines are also called, the dominance of cellphone usage aligns with contemporary communication trends. This results in their rendition of random sampling, which contributes to the external validity of data. However, reaching voters by phone has become more difficult due to low response rates requiring many attempts to secure responses from certain individuals (response rates are as  low as 2% at NYC). In fact, according to an article by Emerson College, telephone polling has become entirely obsolete [@kimball_polling]. Despite this, the phone methodology is still seen as effective for its ability to quickly reach a diverse voter pool across various states as well as seen as the golden standard for polling. Although this can be seen as random sampling, the concern of low response rates and reliance on dialing to reach voters raises concerns of selection bias. There are questions as to whether a certain demographic will be more responsive to being reached through a cell phone. Moreover, voters that are reached over the phone may feel the urge to provide more ‘socially desirable’ responses that are not reflective of their true opinions.

## Non-Response Remedies
With a response rate of 2%, there is valid concern over non-response bias underrepresenting certain demographics. The poll addresses non-response through a combination of front-end sampling adjustments (calling harder-to-reach voters) and post-survey weighting. Weighting is applied to ensure that the final sample aligns with the demographic and political composition of the electorate, based on voter file data. Non-response bias is a persistent challenge, and even though weighting helps, it's impossible to be certain that respondents are fully representative of those who do not respond. Experiments with alternative methods, such as mail surveys with incentives of cash vouchers have shown to increase response rate up to 30% [@nytpoll].

# Sampling Approach: Stratified Sampling

##  Methodology
The Times/Siena Poll employs a stratified random sampling approach to ensure that the sample accurately represents the diversity within the target population of registered voters in the United States. This method involves dividing the population into distinct subgroups, or “strata,” based on key characteristics such as state, age, gender, race, education level, and political affiliation. For example, in presidential elections, particular emphasis is placed on voters from swing states like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on these battleground states, the poll aims to capture the opinions of voters who are most likely to influence the election outcome. Once the strata are established, respondents are randomly selected within each group, ensuring that all demographic groups are appropriately represented in proportion to their actual distribution within the electorate. Furthermore, post-survey weighting is employed to more closely resemble this distribution.

## Advantages
Stratified sampling offers the advantage of higher precision in representing the population's diversity, as each subgroup is proportionally represented. By breaking down the sample into distinct groups, the Times/Siena Poll can adjust for demographic variations between states or voter groups, which is particularly useful in a diverse electorate. This approach helps to minimize sampling error by ensuring that all significant voter subgroups are included, thus increasing the reliability of the poll’s predictions. It is standard practice for them to sample 1,000 respondents for their polls [@nytpoll]. Using stratified sampling to have this sample size closely resemble that of the entire target population cuts costs and provides efficient real-time data.

## Stratified Sampling over Random Sampling

@@

## Challenges and Trade-offs

@@

## Evaluation of Questionnaire Design in the Times/Siena Poll

### Strengths

@@

### Limitations

@@

# Appendix B {-}

@@

###   Referral Mechanics
To further broaden our reach without exponentially increasing advertising costs, we incorporate a referral system that rewards respondents for inviting others to complete the survey [@tellingstories]. Each referral allows the original respondent to gain an additional entry in the lottery, increasing their chances of winning. This referral incentive taps into snowball sampling, where respondents refer people within their network to participate, expanding our pool of respondents organically.
However, using snowball sampling in an election survey introduces potential biases, particularly the risk of overrepresentation of individuals with similar political views, as respondents are likely to refer to close contacts who may share similar ideologies. To mitigate this, we will place specific restrictions on the referral program. Each respondent will be limited to referring only one other individual, minimizing the potential for overly clustered referrals. Additionally, we will require respondents to identify their relationship with each referral, helping us monitor homogeneity in our responses. By balancing the benefits of snowball sampling with these restrictions, we aim to expand the reach of our survey into difficult to reach demographics without compromising the representativeness of the sample.

# Budget Allocation
  Advertisements for sampling recruitment: $70,000
  Prize Pool for Incentive: $5,000
  Data Processing and Cleaning: $15,000
  Miscellaneous Expenses: $5,000
  Petty Cash: $5,000

#  Idealized Questionnaire and Survey

@@

# Model details {#sec-model-details}

## Posterior predictive check

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
pp_check(model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

#| echo: false
#| warning: false
#| message: false
posterior_predictions <- posterior_predict(model, draws = 100)
ppc_dens_overlay(y = data_filtered$average_pct/100, yrep = posterior_predictions) +
  labs(title = "Posterior Predictive Check",
       x = "Rate of support",
       y = "Density")
```

## Diagnostics

>>>>

---
title: "Forecast of the 2024 Presidential Election through the Lens of Turbulent Swing States"
subtitle: "Kamala Harris Leads the Race With 276 Electoral Votes With Support in 4 of 7 Swing States"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/president-polls)."
date: today
date-format: long
abstract: "The 2024 U.S. presidential election is shaping up to be a highly swing race, with polling data playing a crucial role in gauging public sentiment and forecasting the potential outcome. This paper builds a predictive model using a poll-of-polls approach, aggregating data from multiple pollsters to forecast the winner of the upcoming election. The analysis incorporates polling data from various sources, filtered to include only high-quality pollsters with a numeric grade above 3, to ensure the reliability of the predictions. By incorporating the poll-of-polls approach with our model, we predict that Kamla Harris edges our Trump in the presidential election. Our model forecasts that with a 1%-4% margin lead in Pennsylvania, Wisconsin, Nevada, and Michigan, Kamala Harris will accumulate 276 electoral votes to win the election. Additionally, a detailed examination of one selected pollster’s methodology is conducted, highlighting strengths, weaknesses, and biases in the survey process. The paper concludes with an idealized polling methodology, designed within a $100K budget, to enhance the accuracy of election forecasts. Results are presented in the context of understanding polling trends, potential biases, and the evolving landscape of voter preferences as the election approaches."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{pdfpages}
---

@@

# Appendix A{-}

## Sampling Methodology Employed by New York Times and Siena College {#sec-Methodology}

@@

### Target Population and Sampling Frame
The poll targets registered voters in the United States, with a particular focus on voters in swing states for presidential races. Consequently, the target population is registered voters of the United States of America and the sampling frame is the voter file, a national list of registered voters from which NYC employs random sampling to create their pollsters [@nytpoll]. This voter file contains information on more than 200 million Americans that includes information about their party affiliation, voting history, and  basic demographic information. From this sampling frame, respondents are randomly selected while balancing key demographic and political characteristics through both front-end sampling and post-survey weighting pursuant to appropriate representation of the entire target population [@nytpoll].

### Recruitment Methods
The poll relies on random digit dialing to reach voters, predominantly via cell phones. While landlines are also called, the dominance of cellphone usage aligns with contemporary communication trends. This results in their rendition of random sampling, which contributes to the external validity of data. However, reaching voters by phone has become more difficult due to low response rates requiring many attempts to secure responses from certain individuals (response rates are as  low as 2% at NYC). In fact, according to an article by Emerson College, telephone polling has become entirely obsolete [@kimball_polling]. Despite this, the phone methodology is still seen as effective for its ability to quickly reach a diverse voter pool across various states as well as seen as the golden standard for polling. Although this can be seen as random sampling, the concern of low response rates and reliance on dialing to reach voters raises concerns of selection bias. There are questions as to whether a certain demographic will be more responsive to being reached through a cell phone. Moreover, voters that are reached over the phone may feel the urge to provide more ‘socially desirable’ responses that are not reflective of their true opinions.

### Non-Response Remedies
With a response rate of 2%, there is valid concern over non-response bias underrepresenting certain demographics. The poll addresses non-response through a combination of front-end sampling adjustments (calling harder-to-reach voters) and post-survey weighting. Weighting is applied to ensure that the final sample aligns with the demographic and political composition of the electorate, based on voter file data. Non-response bias is a persistent challenge, and even though weighting helps, it's impossible to be certain that respondents are fully representative of those who do not respond. Experiments with alternative methods, such as mail surveys with incentives of cash vouchers have shown to increase response rate up to 30% [@nytpoll].

## Sampling Approach: Stratified Sampling {#sec-stratified-sampling}

###  Methodology
The Times/Siena Poll employs a stratified random sampling approach to ensure that the sample accurately represents the diversity within the target population of registered voters in the United States. This method involves dividing the population into distinct subgroups, or “strata,” based on key characteristics such as state, age, gender, race, education level, and political affiliation. For example, in presidential elections, particular emphasis is placed on voters from swing states like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on these battleground states, the poll aims to capture the opinions of voters who are most likely to influence the election outcome. Once the strata are established, respondents are randomly selected within each group, ensuring that all demographic groups are appropriately represented in proportion to their actual distribution within the electorate. Furthermore, post-survey weighting is employed to more closely resemble this distribution.

### Advantages
Stratified sampling offers the advantage of higher precision in representing the population's diversity, as each subgroup is proportionally represented. By breaking down the sample into distinct groups, the Times/Siena Poll can adjust for demographic variations between states or voter groups, which is particularly useful in a diverse electorate. This approach helps to minimize sampling error by ensuring that all significant voter subgroups are included, thus increasing the reliability of the poll’s predictions. It is standard practice for them to sample 1,000 respondents for their polls [@nytpoll]. Using stratified sampling to have this sample size closely resemble that of the entire target population cuts costs and provides efficient real-time data.

### Stratified Sampling over Random Sampling

@@

### Challenges and Trade-offs

@@

### Evaluation of Questionnaire Design in the Times/Siena Poll

#### Strengths

@@

#### Limitations

@@

# Appendix B{-}

@@

###   Referral Mechanics
To further broaden our reach without exponentially increasing advertising costs, we incorporate a referral system that rewards respondents for inviting others to complete the survey [@tellingstories]. Each referral allows the original respondent to gain an additional entry in the lottery, increasing their chances of winning. This referral incentive taps into snowball sampling, where respondents refer people within their network to participate, expanding our pool of respondents organically.

However, using snowball sampling in an election survey introduces potential biases, particularly the risk of overrepresentation of individuals with similar political views, as respondents are likely to refer to close contacts who may share similar ideologies. To mitigate this, we will place specific restrictions on the referral program. Each respondent will be limited to referring only one other individual, minimizing the potential for overly clustered referrals. Additionally, we will require respondents to identify their relationship with each referral, helping us monitor homogeneity in our responses. By balancing the benefits of snowball sampling with these restrictions, we aim to expand the reach of our survey into difficult to reach demographics without compromising the representativeness of the sample.

## Budget Allocation
  Advertisements for sampling recruitment: $70,000
  Prize Pool for Incentive: $5,000
  Data Processing and Cleaning: $15,000
  Miscellaneous Expenses: $5,000
  Petty Cash: $5,000

## Idealized Questionnaire and Survey

@@

## Model details {#sec-model-details}

### Posterior predictive check

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
pp_check(model) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(title = "Posterior Predictive Check",
       x = "Rate of support",
       y = "Density")

posterior_vs_prior(model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

### Diagnostics
